\input{/Users/joshyv/Research/misc/latex_paper} 
\title{Proof Corrections}
\author{Joshua Vogelstein}

\begin{document}
\maketitle
Items are enumerated as in the faxed document.

\section{Main article errors}

\begin{enumerate}
\item Remove the word ``the''
\item Replace``significant recent progress'' with ``significant progress recently''
\item Replace ``Earlier this year'' with ``Last year''
\item Two comments on Equation (2)
\ita 
\item Break this equation (into multiple lines) at the $=$ 
\item Replace $\varepsilon_t$ with $\varepsilon_{c,t}$
\itb
\item Replace ``$t$th'' with ``$t$-th''
\item It seems as if in $\Ca_b$, the ``$b$'' is bold.  It should not be.
\item Replace ``referred to as'' with ``called''
\item Replace ``i)'' and ``ii)'' with ``(i)'' and ``(ii)'', respectively 
\item Replace $\frac{def}{=}$ with $\overset{def}{=}$, ie, the ``$=$'' should be in line with the rest of the equation, and the ``def'' should be just above it in a smaller font.  This is a recurring problem.  \label{def}
\item Same as \ref{def}
\item Multiple comments on Equation (6)
\ita
\item Same as \ref{def}
\item Replace stuff after $=$ with
\begin{align*}
\begin{cases}
\mathcal{N}\big(\Ca_t; \widehat{\mu}(n_t), \sigma_c^2 \Delta\big) (p \Delta)& \text{if } n_t=1 \\
\mathcal{N}\big(\Ca_t; \widehat{\mu}(n_t), \sigma_c^2 \Delta\big) (1-p \Delta)& \text{otherwise,}\\
\end{cases}
\end{align*}
\item Try to fit within a single column, given the above substitution
\itb
\item Replace ``Now the goal is to efficiently estimate $\p(\ve{H}_t | \ve{O}_{1:T})=\p(n_t, \Ca_t \mid F_{1:T})$  for all $t$, the posterior distribution of the hidden signals, given all the observations.'' with `` Now the goal is to efficiently estimate $\p(\ve{H}_t | \ve{O}_{1:T})=\p(n_t, \Ca_t \mid F_{1:T})$, the posterior distribution of the hidden signals, given all the observations, for all $t$. ''
\item Replace ``$P_{\bth (\bH_t | \bO_{1:t})}$'' with ``$\p(\bH_t | \bO_{1:t})$'', ie, only the $\bth$ gets subscripted.
\item In Equation (9), after the final $=$, it should read: $$  \p \big( F_t \mid \Ca_t^{(i)} \big) \p \big( \Ca_t^{(i)}   \mid \Ca^{(i)}_{t-1}, n_t^{(i)} \big)  \p \big( n_t^{(i)} \big)/Z, $$  The only acceptable places to put line breaks are between the ``$)$'''s and ``$P$'''s.
\item Replace ``$\{n_t,\Ca_t\}$'' with ``$n_t$ or $\Ca_t$''
\item Two remakrs about Equation (12)
\ita
\item Replace $$\frac{\argmax}{\bth}$$ with $$\argmax_{\bth}$$ (note that argmax goes in line with the rest of the text, and $\bth$ goes below it in a smaller font, similar to \ref{def}).
\item Note that the $\bth$ should be in bold.
\item The ``,'' should be after the final ``$)$'', not before
\itb
\item The ``$\theta'$'' should be ``$\bth'$'', ie, bold.
\item Replace ``variance'' with ``noise term''
\item Remove ``Main Result.'' 
\item Remove sentence ``Note that\ldots of generality.''
\item See \ref{def}
\item Repalce ``($\q=\p(\ve{H}_t^{(i)} | \ve{H}_{t-1}^{(i)}, \ve{O}_t)$)'' with   ``, $\q=\p(\ve{H}_t^{(i)} | \ve{H}_{t-1}^{(i)}, \ve{O}_t)$,''
\item Replace ``($n_{1:T}$)'' with ``, $n_{1:T}$'' and ``($\tau$)'' with ``, $\tau$,''
\item Remove ``Saturated simulation.''
\item Replace ``$\xi=4 \times 10^{-4}$ a.u.'' with ``$\xi=4 \times 10^{-4}$ $\mu$A/photon, and replace $\sig_F=10^{-4}$ a.u.'' with ``$\sig_F=10^{-4}$ $\mu$A''
\item Remove ``In vitro bursts.''
\item Replace ``i)'' with ``(i)'' 
\item Remove ``Real data saturation''
\item Replace and ``ii)'' with ``(ii)''
\item Add the word ``its'' between ``zero with'' and ``time constant''
\item Remove ``Array of inferences.''
\item Remove ``Generalized linear model particle filter smoother.''
\item Remove ``In vitro data superresolution''
\item Replace ``PFS both improves inference accuracy over the optimal linear method and provides'' with ``PFS both (i) improves inference accuracy over the optimal linear method and (ii) provides''
\item Replace ``$\{\alpha,\beta,\sig_F\}$'' with ``$\{\alpha, \beta, \xi, \sig_F\}$''
\item Replace ``i)'' with ``(i)'' and ``ii)'' with ``(ii)''
\end{enumerate}

\section{Appendix errors}

\begin{enumerate}

\item Two things
\ita 
\item See \ref{def}
\item Break at $=$
\itb
\item Break into two lines if possible
\item Try to into two lines
\item Replace ``$\{\Ca_{t-1}\}^{(i)}$'' with ``$\Ca_{t-1}^{(i)}$''
\item Equation (34) should be rewritten as below.  Note the location of the line break
\begin{multline*}
g'(x) = \left(\frac{k_d (\beta -F_t)}{F_t - \beta - \alpha}\right)^{1/n}  \frac{n k_d (\beta - F_t)}{F-\beta-\alpha} \\ 
\left(-\frac{k_d}{F-\beta-\alpha}-\frac{k_d(\beta-F_t)}{(F_t-\beta-\alpha)^2}\right).
\end{multline*}
\item Replace ``are function of all the parameters: $\alpha, \beta, \xi, \sig_F, n$ and $k_d$'' with ``are function of $\{\alpha, \, \beta, \, \xi, \, \sig_F, \, n, \, k_d\}$''
\item Line break after $=$ sign
\item Two things
\ita \item See \ref{def} \item Replace cent sign with $\mathbb{Z}$ \itb
\item Replace ' with $\mathbb{Z}$
\item Remove ``Laplace \ldots distribution.''
\item Replace ``The circles'' with ``The filled circles''
\item Line beginning with $q_{\bth}^S$ should not be indented
\item (10) and (11) should not be links to references, they are sequences
\item Remove ``Sampling strategies:'' and capitalize ``the'' 
\item (10) is a sequence, not a reference, no link
\item Remove ``Mixture approximation:'' and capitalize ``appropriate''
\item ``Computing\ldots'' should be a one level down in the heading hierarchy
\item This equation should not have a line break where the second line break is.  I suggest:
\begin{multline*} 
\p^{NL}(F_v \mid \Ca_{v-1}) = \sum_{n=0,1} a_{n,v-1} \\
\int \p^{NL}(F_v \mid \Ca_v) \p(\Ca_v \mid \Ca_{v-1}, n_v=n) d\Ca_v,
\end{multline*}
or 
\begin{multline*} 
\p^{NL}(F_v \mid \Ca_{v-1}) = \sum_{n=0,1} a_{n,v-1} \int \p^{NL}(F_v \mid \Ca_v) \\
 \p(\Ca_v \mid \Ca_{v-1}, n_v=n) d\Ca_v,
\end{multline*}
\item ``$a_{1, \quad v-1}$'' should be ``$a_{1,v-1}$'', ie, there seems to be a space between the comma and the 'v', which should not be present. same is true for $a_{0,v-1}$.  This problem repeats in several places \label{comma}
\item Indenting is totally inappropriate.  Please use:
\begin{multline*}
\p(\Ca_v \mid \Ca_{v-1}, n_v) =\\
\mathcal{N}(\Ca_v; \Ca_{v-1} -\Delta/\tau (\Ca_{v-1} - \Ca_b) + A n_v, \sigma_c^2 \Delta),
\end{multline*}
\item Line break is again inappropriate.  replace with
\begin{multline*}
\int \p^{NL}(F_v \mid \Ca_v) \p(\Ca_v \mid \Ca_{v-1}, n_v=n) d\Ca_v
\\= \int \mathcal{N}\big(\Ca_v; \widetilde{\mu}_v, \widetilde{\sigma}_v^2\big)  \times \mathcal{N}(\Ca_v; \Ca_{v-1} -\Delta/\tau (\Ca_{v-1} - \Ca_b) + A n_v, \sigma_c^2 \Delta).
\end{multline*}
\item Replace $\chi$ with $C_v$
\item Replace $\chi$ with $C_v$
\item Line break should be:
\begin{multline*}
\p^{NL}(F_v \mid \Ca_{v-1}) = \\ \sum_{n=0,1} a_{n,v-1}  \mathcal{N} \left(\Ca_{v-1}; \widetilde{\mu}^{S}_v(n), (\widetilde{\sigma}_v^S)^2 \right),%\mathcal{N}\left(\Ca_{v-1};\frac{ \widetilde{\mu}_v- A n_v - \frac{\Delta}{\tau} \Ca_b}{1-\frac{\Delta}{\tau}}, \frac{\widetilde{\sigma}_v^2 + \sigma_c^2 \Delta}{(1-\frac{\Delta}{\tau})^2}\right),
\end{multline*}
\item  Line break inappropriate.  Replace with:

\begin{multline*} 
\p^{NL}(F_v \mid \Ca_{t-1}) =\\ \sum_{n=0,1} a_{n,t-1}
\sum_{m=1}^{2^{v-t}} a_{mt} \mathcal{N}\left(\Ca_{t-1}; \widetilde{\mu}^{S}_{mt}(n),(\widetilde{\sigma}^S_t)^2 \right), %= \sum_m^{2^{v-(t-1)}} a_m \mathcal{N}(\Ca_{t-1}; \mu_{mt}, \sigma_t^2)
\end{multline*}
\item ``Superresolution sampling spikes'' should be one level down in hierarchy, ie, it should be indented \label{ind}
\item Replace ${n_t}^{(i)}$ with $n_t^{(i)}$, ie, the $(i)$ should be directly above $t$, not off to the side.  this happens in 3 places on this line, and elsewhere \label{n}
\item ``Superresolution\dots'' should be indented, just as \ref{ind}
\item See \ref{n}
\item See \ref{n} (twice on this line)
\item Line break is inappropriate, move it to after $=$
\item Same as \ref{comma}
\item Same as \ref{comma}
\item Two comments on this equation
\ita
\item Replace $\frac{\argmax}{\{b,\bk,\bomega\}}$ with $$\argmax_{\{b,\bk,\bomega\}}$$ in both instances (ie, argmax goes in line with the rest, $\{b,\bk,\bomega\}$ goes under in a smaller font
\item (61) should be on the bottom line, not the middle one
\itb
\item Replace $\overset{\argmax}{\tau,A,\Ca_b,\sig_c}$ with $$\argmax_{\tau,A,\Ca_b\geq 0}$$.  Note a few differences
\ena
\item argmax is inline with rest of text, the other stuff is below
\item $\sig_c$ has been replaced with $\leq$
\enb
\item Again, argmin should go inline, $x>0$ should go underneath.  equation should look like
\begin{align*} 
\widehat{\ve{x}} = \argmin_{\ve{x} > 0} \frac{1}{2}\ve{x}' \ve{Q} \ve{x} + \ve{L}' \ve{x},
\end{align*}
\item Remove $\times$
\item Same deal again with the argmin, equation should look like:
\begin{equation*}
\begin{split}
\widehat{\ve{x}} &= \frac{1}{2}  \argmin_{\ve{x_p} \geq 0, \forall p} \sum_{t=1}^T \sum_{i,j=1}^N J^{(i,j)}_{t,t-1}  \norm{\ve{C}^{(i,j)}_t \ve{x} + d^{(i,j)}_t}_2^2,
\end{split}
\end{equation*}
\item Again with the argmax, (69a) should look like
\begin{multline*}
\widehat{\sigma_c}^2 = \argmax_{\sigma_c^2} \sum_{t=1}^T \sum_{i,j=1}^N J^{(i,j)}_{t,t-1}  \Bigg( -\frac{1}{2} \ln (2 \pi \sigma_c^2 \Delta) \\- \frac{1}{2} \frac{\big(\Ca_t^{(i)} - \mu_{t,t-1}^{(i,j)}\big)^2}{\sigma_c^2 \Delta}\Bigg)
\end{multline*}
\item Same with argmin, equation should look like
\begin{multline*} 
\{\widehat{\alpha}, \widehat{\beta}\} = \argmin_{\alpha, \beta \geq 0} \sum_{t=1}^T \sum_{i=1}^N  \frac{(F_t-\alpha S(\Ca_t) - \beta)^2}{\xi S(\Ca_t) + \sigma_F} \\ + \ln\big(\xi S(\Ca_t) + \sigma_F\big),
\end{multline*}
\item Another problem with argmin, equation should look like
\begin{align*}
\{\widehat{\xi},\widehat{\sigma_F}\} = \argmin_{\xi, \sigma_F \geq 0} (r_t-\xi S(\Ca_t) - \sigma_F)^2,
\end{align*}
\end{enumerate}


\newpage \clearpage
\section{response to inquiries}

\ena
\item ok.
\item ok.
\item ok.
\item See \cite{YusteKonnerth06}
\item See \cite{TayYue07}
\item See \cite{YasudaSvoboda04}
\item See \cite{BorstAbarbanel07}
\item See \cite{SasakiIkegaya08}
\item See \cite{DoucetGordon01}
\item See \cite{DempsterRubin77}
\item See \cite{SamejimaKimura04}
\item This is correct.  NIPS is a conference proceedings.
\item Please replace with \cite{HuysPaninski09}
\item Maybe also look for Transactions of the ASME, as in \cite{Kalman60}
\item Please replace with \cite{HuysPaninski09}
\item Journal title is ``Network: Computation in Neural Systems'', as suggested by \cite{Paninski04}
\item See \cite{VogelsteinPaninski09a}  
\item See \cite{PillowSimoncelli08}
\item See \cite{ShumwayStoffer06}
\item Replace ``To sample conditioned on the next observation'' with ``To sample $\{n_t, \Ca_t\}$ conditioned on the next observation''
\enb


\clearpage \newpage
\bibliography{/Users/joshyv/Research/misc/biblist}
\addcontentsline{toc}{section}{References}
\bibliographystyle{apalike}
%\bibliographystyle{biophysj}


\end{document}
