\input{/Users/joshyv/Research/misc/latex_paper} 
\title{Proof Corrections}
\author{Joshua Vogelstein}

\begin{document}
\maketitle
Items are enumerated as in the faxed document.

\begin{enumerate}

\item Remove the word ``the''
\item Replace``significant recent progress'' with ``significant progress recently''
\item Replace ``Earlier this year'' with ``Last year''
\item Two comments on Equation (2)
\ita 
\item Break this equation (into multiple lines) at the $=$ 
\item Replace $\varepsilon_t$ with $\varepsilon_{c,t}$
\itb
\item Replace ``$t$th'' with ``$t$-th''
\item It seems as if in $\Ca_b$, the ``$b$'' is bold.  It should not be.
\item Replace ``referred to as'' with ``called''
\item Replace ``i)'' and ``ii)'' with ``(i)'' and ``(ii)'', respectively 
\item Replace $\frac{def}{=}$ with $\overset{def}{=}$, ie, the ``$=$'' should be in line with the rest of the equation, and the ``def'' should be just above it in a smaller font.  This is a recurring problem.  \label{def}
\item Same as \ref{def}
\item Multiple comments on Equation (6)
\ita
\item Same as \ref{def}
\item Replace stuff after $=$ with
\begin{align}
\begin{cases}
\mathcal{N}\big(\Ca_t; \widehat{\mu}(n_t), \sigma_c^2 \Delta\big) (p \Delta)& \text{if } n_t=1 \\
\mathcal{N}\big(\Ca_t; \widehat{\mu}(n_t), \sigma_c^2 \Delta\big) (1-p \Delta)& \text{otherwise,}\\
\end{cases}
\end{align}
\item Try to fit within a single column, given the above substitution
\itb
\item Replace ``Now the goal is to efficiently estimate $\p(\ve{H}_t | \ve{O}_{1:T})=\p(n_t, \Ca_t \mid F_{1:T})$  for all $t$, the posterior distribution of the hidden signals, given all the observations.'' with `` Now the goal is to efficiently estimate $\p(\ve{H}_t | \ve{O}_{1:T})=\p(n_t, \Ca_t \mid F_{1:T})$, the posterior distribution of the hidden signals, given all the observations, for all $t$. ''
\item Replace ``$P_{\bth (\bH_t | \bO_{1:t})}$'' with ``$\p(\bH_t | \bO_{1:t})$'', ie, only the $\bth$ gets subscripted.
\item In Equation (9), after the final $=$, it should read: $$  \p \big( F_t \mid \Ca_t^{(i)} \big) \p \big( \Ca_t^{(i)}   \mid \Ca^{(i)}_{t-1}, n_t^{(i)} \big)  \p \big( n_t^{(i)} \big)/Z, $$  The only acceptable places to put line breaks are between the ``$)$'''s and ``$P$'''s.
\item Replace ``$\{n_t,\Ca_t\}$'' with ``$n_t$ or $\Ca_t$''
\item Two remakrs about Equation (12)
\ita
\item Replace $$\frac{\argmax}{\bth}$$ with $$\argmax_{\bth}$$ (note that argmax goes in line with the rest of the text, and $\bth$ goes below it in a smaller font, similar to \ref{def}).
\item Note that the $\bth$ should be in bold.
\item The ``,'' should be after the final ``$)$'', not before
\itb
\item The ``$\theta'$'' should be ``$\bth'$'', ie, bold.
\item Replace ``variance'' with ``noise term''
\item Remove ``Main Result.'' 
\item Remove sentence ``Note that\ldots of generality.''
\item See \ref{def}
\item Repalce ``($\q=\p(\ve{H}_t^{(i)} | \ve{H}_{t-1}^{(i)}, \ve{O}_t)$)'' with   ``, $\q=\p(\ve{H}_t^{(i)} | \ve{H}_{t-1}^{(i)}, \ve{O}_t)$,''
\item Replace ``($n_{1:T}$)'' with ``, $n_{1:T}$'' and ``($\tau$)'' with ``, $\tau$,''
\item Remove ``Saturated simulation.''
\end{enumerate}


















\end{document}
